---
title: "cutoffvalue: Clean your data"
author: "Lea R Medeiros"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
vignette: >
  %\VignetteIndexEntry{cutoffvalue: Clean your data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>"
)
library(cutoffvalue)
library(readxl)
```

## Using the `cleandata` function to clean up your raw dataset
This vignette will only go over the cleandata function and assumes you have installed the cutoffvalue package and loaded the necessary packages (see the Getting Started vignette for more information). If you need help on other topics, please see either the main vignette (Help) or refer to the vignette for another topic.

This function is used to clean up your dataset by removing any blank rows and transforms the data into a list of numbers. This list of values (`mydata$data`) is returned to the Environment along with the minimum (`mydata$lower`) and maximum values (`mydata$upper`), which are used in subsequent functions. **The mydata label must be applied to this function for it to be used in subsequent functions.**

> If nothing is specified (i.e., only "cleandata()" is typed into the console), this function will use an internal dataset object identified as "cutoffvalue:::exampledata" as the default dataset. 

```{r cleandata}
mydata <- cleandata(cutoffvalue:::exampledata)
```
