---
title: "cutoffvalue: Determine parameters to generate curves and a graph"
author: "Lea R Medeiros"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
vignette: >
  %\VignetteIndexEntry{cutoffvalue: Determine parameters to generate curves and a graph}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>"
)
library(cutoffvalue)
library(readxl)
```

## Using the `fitparams` function to produce a basic histogram and associated parameters

The `fitparams` function will produce a basic histogram from the dataset, which is then used to generate certain parameters for the curve fitting functions in subsequent functions. It assumes you have installed the cutoffvalue package and loaded the necessary packages (see the Getting Started vignette for more information). If you need help on other topics, please see either the main vignette (Help) or refer to the vignette for another topic. 

> If nothing is specified (i.e., only "fitparams()" is typed into the console), this function will use an internal dataset object identified as "cutoffvalue:::exampledata" as the default dataset. 

The `fitparams` function returns a list of values to the Environment that are used in subsequent functions. As such, you should alter the number of breaks to produce a graph representative of what you would like to see in the final plot - if not specified, the default is 15. 

```{r basic-histogram, results='hide', fig.keep='last', fig.align='center', dpi=150, fig.dim=c(9, 5), out.width=700, out.height=390}
fit <- fitparams(cutoffvalue:::exampledata, breaks = 15)
```

_This histogram should be a general outline of what you would like the histogram in the final plot to look like. If it is not, change the number of breaks until it is. The scale of the axes in the final graph will better represent your dataset, so don't worry about those._
