---
title: "cutoffvalue: Overview & Getting Started"
author: "Lea R Medeiros"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
vignette: >
  %\VignetteIndexEntry{cutoffvalue: Overview & Getting Started}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup-chunk, include = FALSE}
knitr::opts_chunk$set(
 collapse = TRUE,
 fig.align = 'center', fig.height = 8, fig.width = 10,
 fig.path='man/cutoffvalue_figures/README-',
 dpi = 300,
 dev = c('jpeg', 'pdf', 'png'),
 echo = TRUE
 )
```

# cutoffvalue

cutoffvalue is a simple R package that implements an updated version of the method first developed and used in Medeiros et al. (2018)[^1]. It can be used to determine an objective cutoff value between a significantly bimodal distribution of log-transformed data and plot a representative graph of the results.

The overall goals of this package are to (1) determine a cutoff value between the upper and lower modes of the dataset and (2) produce a nice graph of the results that includes a histogram of the data, the two models fit to the upper and lower modes, and a line depicting the cutoff value. The functions are written to be run independently, so that only two functions need to be run to get the necessary information:

- modes() <- Determines modality of the dataset, which is an assumption for the remaining functions
- cutoffplot() <- Determines the cutoff value and produces a nice graph of the results

The functions in this package are written to utilize the example data set (an internal dataset object identified as "exampledata") and will use it by default if a dataset is not provided. Below are the instructions on how to install the cutoffvalue package, setup the R Studio environment, and suggestions on how to import and convert your dataset into the proper format. Please see the vignette for more information about each function.

## Installation of the cutoffvalue package

You can install the development version of cutoffvalue from [GitHub](https://github.com/lea-medeiros/cutoffvalue.git) with:

```{r install-package}
# devtools::install_github("lea-medeiros/cutoffvalue", dependencies = TRUE, build_vignettes = TRUE)
```


## Setup RStudio to use the cutoffvalue package

Load the packages that you will need.

```{r load-packages, message=FALSE}
library(cutoffvalue)
library(readxl) # Only necessary if you will be importing an excel file (see below)
```
