---
title: "cutoffvalue: An R Package for Cutoff Value Determination & Graphing"
author: "Lea R Medeiros"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
vignette: >
  %\VignetteIndexEntry{cutoffvalue: An R Package for Cutoff Value Determination & Graphing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = TRUE,
  comment = "#>"
)
```

# Overview of the cutoffvalue package
## Description

cutoffvalue is a simple R package that implements an updated version of the method first developed and used in Medeiros et al. (2018)^[Medeiros LR, Galbreath PF, Knudsen CM, Stockton CA, Koch IJ, Bosch WJ, Narum SR, Nagler JJ, Pierce AL (2018) Plasma 11-Ketotestosterone in Individual Age-1 Spring Chinook Salmon Males Accurately Predicts Age-2 Maturation Status. Transactions of the American Fisheries Society 147 (6):1042-1051. doi:10.1002/tafs.10097]. It can be used to determine an objective cutoff value between a significantly bimodal distribution of log-transformed data and plot a representative graph of the results.

The functions in this package are written to utilize the example data set (an internal dataset object identified by "rawdata") and will use it by default if the path for a dataset is not provided. The examples below specify this, along with any other default parameters for each function.

## Vignette Info & Suggested Workflow

This vignette will go over the various functions included in cutoffvalue using the included `example data`. The overall goals of this package are to (1) determine a cutoff value between the upper and lower modes of the dataset and (2) produce a nice graph of the results that includes a histogram of the data, the two models fit to the upper and lower modes, and a line depicting the cutoff value. The functions are written to be run independently, so that only two functions need to be run to get the necessary information:

- modes() <- Determines modality of the dataset, which is an assumption for the remaining functions
- cutoffplot() <- Determines the cutoff value and produces a nice graph of the results

This vignette is written in accordance with the suggested workflow using these two functions. Subsequent sections discuss the functions that are included within them.

## Installation of the package

You can install the latest version of cutoffvalue from [GitHub](https://github.com/lea-medeiros/cutoffvalue.git) with:

``` r
# install.packages("devtools")
devtools::install_github("lea-medeiros/cutoffvalue", dependencies = TRUE, build_vignettes = TRUE)
```

## Setup R Studio

Load the cutoffvalue package
```{r load-packages, message=FALSE}
library(cutoffvalue)
```

## Import your raw dataset

Import the data file to be used in the analyses and graph. The package includes a dataset for use as an example - this object is accessible as "rawdata" and will be used in the examples.

Import your dataset any way you prefer. However, keep in mind that data should be organized as a single column of log- or natural log-transformed data. I find that the easiest way to import data is to use the "Import Dataset" function built into R Studio, but you can also use code (an example is provided below).

``` r
library(readxl)
yourrawdata <- read_excel("/path/to/your/excel/data")
```

Each function in this package uses the provided dataset (whether it's the example dataset or one you provided) and cleans it up (via the `cleandata` function, see below) to remove any blank cells. This function then provides a list of objects: the data (`mydata$data`), the maximum value (`mydata$upper`), and the minimum value (`mydata$lower`), all of which are then used in subsequent calculations.

<br/><br/>

# Using two functions to determine modality and generate the final plot

## Use the `modes` function to determine modality

Evaluating an objective and valid cutoff value depends upon whether the dataset is bimodal. Which means that the modality of the dataset should be determined before proceeding. The null hypothesis of this test is that the dataset is unimodal; an excess mass statistic associated with a p-value of less than 0.05 implies it is more than unimodal. It is strongly suggested that you run this test prior to any other function in cutoffvalue; however, how you proceed depends on your knowledge and understanding of the data.

> The `modes` function uses the example dataset (an internal dataset object identified by "rawdata") by default. Thus, if the path for the dataset is not specified (e.g., running modes() in the console), this is the data that will be used.

> When given a label (i.e., "modetest" in the example below), this function returns the Excess Mass statistic and associated p-value to the Environment.

Please keep in mind that the default parameters of cutoffvalue assume a bimodal distribution - any other modalities may cause inaccuracies in the results.

```{r modetest, warning=FALSE, echo=1}
modetest <- modes(rawdata)
```

_The `modes()` function returns the p-value and excess mass statistic along with instructions on how to proceed based on the p-value. If the p-value is less than 0.05, accept the alternative hypothesis (data is at least bimodal) and proceed with analysis. However, if the p-value is more than 0.05, the data is unimodal and the following analyses are not entirely valid._

## Use the `cutoffplot` function to plot the final graph

Plot a pretty graph that includes a histogram of the data, curve lines for each mode generated from the model results, the cutoff value depicted as a line, and labels customized for your dataset. 

> The `cutoffplot` function uses the example dataset (an internal dataset object identified by "rawdata") by default. Thus, if the path for the dataset is not specified (e.g., running cutoffplot() in the console), this is the data that will be used. Parameters (and their defaults) for the resulting graph are listed below.

> When given a label (i.e., "plotty" in the example below), this function returns the log-transformed cutoff value to the Environment. 

### Specify graph labels

You will likely want to change the following parameters to match your own dataset and preferences. If nothing is specified in the `cutoffplot` function, then these are the defaults that will be used in the graph.

```{r graph-parameters, echo=TRUE}
title <- "Plasma 11-KT levels in age-2 male spring chinook"  # Graph title
xlab <- "Plasma [11-KT] (ng/mL)" # X-axis label
cutofflab <- "Minijack cutoff" # label for cutoff value on graph
cutoffunits <- "(ng/mL)" # units for cutoff value
LowerMode_col <- "red" # line color for the lower mode
LowerMode_lty <- 1 # line type for the lower mode
LowerMode_lwd <- 2 # line width for the lower mode
UpperMode_col <- "purple" # line color for the upper mode
UpperMode_lty <- 1 # line type for the upper mode
UpperMode_lwd <- 2 # line width for the upper mode
cutoffvalue_col <- "black" # line color for the cutoff value
cutoffvalue_lty <- 2 # line type for the cutoff value
cutoffvalue_lwd <- 2 # line width for the cutoff value
```

### Plot the graph
```{r graph, results='hide', fig.keep='last', fig.align='center', dpi=150, fig.dim=c(9, 5), out.width=700, out.height=390}
plottyMcplotty <- cutoffplot(rawdata, title, xlab, cutofflab, cutoffunits, LowerMode_col, LowerMode_lty, LowerMode_lwd, UpperMode_col, UpperMode_lty, UpperMode_lwd, cutoffvalue_col, cutoffvalue_lty, cutoffvalue_lwd)
```

<br/><br/>

# Other Functions Included in the cutoffvalue Package

The following functions are included in the cutoffvalue package and can be run independently. They are discussed in order of operations for having the `cutoffplot` function work properly.

## Use the `cleandata` function to import your raw dataset

This function is used to clean up your dataset by removing any blank rows and transforms the data into a list of numbers. This list of values (`mydata$data`) is returned to the Environment along with the minimum (`mydata$lower`) and maximum values (`mydata$upper`), which are used in subsequent functions. **The mydata label must be applied to this function for it to be used in subsequent functions.**

> If nothing is specified (i.e., only "cleandata()" is typed into the console), this function will use an internal dataset object identified by "rawdata" as the default dataset. 

```{r cleandata}
mydata <- cleandata(rawdata)
```

## Use the `datamodel` function to generate models for each mode of the dataset

The `datamodel` function fits two component mixture models to the data and plots a rough histogram with the fitted lines. It also defines the index.lower value to be used in the find.cutoff function. 

> The `datamodel` function uses the example dataset (an internal dataset object identified by "rawdata") by default. Thus, if the path for the dataset is not specified (e.g., running importdata() in the console), this is the dataset that will be used.

> When given a label (i.e., "model" in the example below), the `datamodel()` function returns a list of 2 objects to the Environment - `model$mydata` and `model$indexLower`, which are used in subsequent functions.

```{r model-data, results='hide', fig.keep='last', fig.align='center', dpi=150, fig.dim=c(9, 5), out.width=700, out.height=390}
model <- datamodel(rawdata)
```

_This isn't the final graph, but still should be inspected to ensure that things look right. In particular, make sure that the point where the two curves intersect is where you are expecting the cutoff to be._

## Use the `findcutoff` function to determine the cutoff value

The `findcutoff` function determines the cutoff value between the two modes with an equal chance of being drawn from either mode. The default probability is set to 50% (i.e., "proba=0.5" in the example below), but the probability can be changed in the code. 

> The defaults for this function are an internal dataset object identified by "rawdata" as the raw dataset and 0.5 as the probability (i.e., running "findcutoff()" in the console will use these values)

> Running the `findcutoff` function with a label (e.g., "cutoff" in the example below) will return the cutoff value to the Environment, but otherwise it does not report to the console. Use "returnValue(cutoff)" if you would like to see the value in the console.

```{r find-cutoff, fig.show='hide'}
cutoff <- findcutoff(rawdata, proba=0.5)
```

_The uniroot lower (`mydata$lower`) and upper values (`mydata$upper`) are determined using the range of "mydata" and will reflect the dataset being analyzed. If there are errors due to the uniroot, consider editing the custom values to something that more generally reflects the range of the data being analyzed._

## Use the `fitparams` function to produce a basic histogram and associated parameters

The `fitparams` function will produce a basic histogram from the dataset, which is then used to generate certain parameters for the curve fitting functions in subsequent functions. As such, you should alter the number of breaks to produce a graph representative of what you would like to see in the final plot - if not specified, the default is 15. 

> If nothing is specified (i.e., only "fitparams()" is typed into the console), this function will use an internal dataset object identified by "rawdata" as the default dataset. 

The `fitparams` function returns a list of values to the Environment that are used in subsequent functions.

```{r basic-histogram, results='hide', fig.keep='last', fig.align='center', dpi=150, fig.dim=c(9, 5), out.width=700, out.height=390}
fit <- fitparams(rawdata, breaks = 15)
```

_This histogram should be a general outline of what you would like the histogram in the final plot to look like. If it is not, change the number of breaks until it is. The scale of the axes in the final graph will better represent your dataset, so don't worry about those._

## Use the `curves` function to generate points for the curves

The `curves` function determines x and y values to calculate the points for the curves that represent the generated models in the final plot. 

> If nothing is specified (i.e., only "curves()" is typed into the console), this function will use an internal dataset object identified by "rawdata" as the default dataset. 

The `curves` function returns a list of 3 objects to the Environment, which are used by the `cutoffplot` function when producing the final plot.

```{r curves, results='hide', fig.show='hide'}
curves <- curves(rawdata)
```
